---
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
#| label: libraries
library(gmodels)
library(dplyr)
library(skimr)
library(gt)
library(ggplot2)
library(GGally)
library(gridExtra)
library(GGally)
library(class)
library(MASS)
library(caret) # for confusion matrix
library(e1071)
library(ROCR) # For AUC
library(corrplot)
library(pROC)      # for ROC plot
library(rpart)
library(randomForest)
library(rpart.plot)
library(knitr)
```

# Introduction {#sec-Intro}

The study aims to predict the factors that impact the subscription of
term deposit (yes/no). The data set contains 10,000 records from a
Portuguese banking institution's directed marketing campaigns conducted
via phone calls. These campaigns targeted clients to promote bank term
deposits. The data set comprises information from 17 campaigns spanning
May 2008 to November 2010.

```{r}
df <- read.csv("group_14.csv") #Import the data set#
```

# Data Pre-processing {#sec-DW}

The process below describes the detailed data wrangling techniques that
are used to get the desired data set. On the further examination of the
data set, it was observed that there were some empty cells present.
These cells were substituted by NA. The categorical columns were
converted to factor type.

```{r}
#Check for missing values#
missing_values <- colSums(is.na(df)) # no missing values
df[df == ""] <- NA #However, empty cells were observed & NA values were assigned
df <- df %>% #Change character variables to factors
  mutate_at(vars(job, marital, education, default, housing, 
                 loan, contact, month, day_of_week, poutcome, y), as.factor)
df_numeric <- df %>%
  select_if(is.numeric) # numeric data set as 'df_numeric'
df_not_numeric <- df %>%
  dplyr::select(job, marital, education, default, housing, loan, contact, month, day_of_week, poutcome, y)
df_new <- cbind(df_not_numeric, df_numeric) # non-numeric data set 'df_not_numéric'
#Summary Statistics#
summary_df_new <- summary(df_new) 
```

From the summary statistics, number of missing values were observed in
categorical columns. These missing values were imputed by Mode.

```{r}
#Impute NA values#
#JOB
sum_job <- summary(df_new$job) # 84 NA values
job_mode <- names(which.max(table(df_new$job))) # "admin."
df_new$job[is.na(df_new$job)] <- job_mode

#MARITAL
sum_mar <- summary(df_new$marital) # 25 NA values
marital_mode <- names(which.max(table(df_new$marital))) # "married"
df_new$marital[is.na(df_new$marital)] <- marital_mode

#EDUCATION
sum_edu <- summary(df_new$education) # 402 NA
edu_mode <- names(which.max(table(df_new$education))) # "university.degree"
df_new$education[is.na(df_new$education)] <- edu_mode

#DEFAULT
dum_def <- summary(df_new$default) # 2151 NA
default_mode <- names(which.max(table(df_new$default)))  # "no"
df_new$default[is.na(df_new$default)] <- default_mode

#HOUSING
sum_hous <- summary(df_new$housing) # 241 NA
housing_mode <- names(which.max(table(df_new$housing))) # "yes"
df_new$housing[is.na(df_new$housing)] <- housing_mode

#LOAN
sum_loan <- summary(df_new$loan) # 241 NA
loan_mode <- names(which.max(table(df_new$loan))) # "no"
df_new$loan[is.na(df_new$loan)] <- loan_mode

#Check for NA values using colSums(is.na(df_new)) resulted in No more NA values
```

The column 'default' has 2151 missing values. Since there are replaced
by the mode, there are 9999 No and 1 yes. This imputation does not
provide any relevant information. Hence, 'default' will be omitted.

```{r}
#| label: fig-dow-age
#| fig-cap: Understandind the Data structure of Day of the Week and Age
#| fig-subcap:
#|   - "Historgram of day_of_week"
#|   - "Boxplot of age"
#| layout-ncol: 2
#| column: page
#| fig-align: left
#| fig-width: 3
#| fig-height: 3

#day-of the week#
dow_dist <- df_new %>% 
  group_by(day_of_week) %>%
  dplyr::summarise(dow_count= n()) 
ggplot(df_new, aes(x = day_of_week, fill = y)) +
  geom_bar(position = "dodge") +
  labs(title = "Dodged barplot",
       x = "Day of Week",
       y = "Count",
       fill = "y")+
  theme_bw()
chi_dow <- chisq.test(df_new$day_of_week, df_new$y) # p-value = 0.4149 > 0.05
# This suggests that there is not enough evidence to conclude a significant association between the two categorical variables.

# Duration : discard as per the data description instruction
# Columns removed #
df_new <- df_new %>%
  dplyr::select(-default, -day_of_week, -duration)
# age 
ggplot(data = df_new, aes(x = "", y = age)) +
  geom_boxplot(fill = "skyblue", color = "black")+
  theme_bw()
df_new$age <- log(df_new$age)
# pdays
summary(df_new$pdays)
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 0.0   999.0   999.0   963.5   999.0   999.0 
```

From the above graph it can be seen that there are the days of the week
do not affect the distribution of the response variable greatly. Due to
this, 'day_of_week' is omitted. This can further be verified by using
chi-squared test of independence. From the summary statistics as well as
the box plot, 'age' seems to display a large variation and great number
of outliers. Logarithmic transformation is applied for robust analysis.
For column 'pdays' it can be observed that is highly skewed, with most
values concentrated in a specific range around '999'. This can reduce
the effectiveness of the variable in splitting the data into homogeneous
groups, which is essential for some classification models. Considering
this, this column is not included henceforth.Binary variable is allotted
to the response variable 'y' .

```{r}
df_new_numeric <- df_new %>%
  select_if(is.numeric)
df_new_not_numeric <- df_new %>%
  dplyr::select(c(job, marital, education, housing,loan, contact, month, poutcome))
y <- df_new$y # Binary Variable for y 
df_new$y_coded <- ifelse(df_new$y == "yes", 1, 0)
y_coded <- df_new$y_coded
data <- cbind(y, y_coded, df_new_not_numeric , df_new_numeric)
```

# Exploratory Data Analysis {#sec-EDA}

The data frame is divided into a training set, validation set and test
set using 50%-25%-25% division. The Exploratory data analysis is
performed on the training data set.

```{r}
set.seed(1)
n <- nrow(data)
ind1 <- sample(c(1:n), floor(0.5*n))
ind2 <- sample(c(1:n)[-ind1], floor(0.25*n))
ind3 <- setdiff(c(1:n),c(ind1,ind2))
data.train <- data[ind1,] # 5000
data.valid <- data[ind2,] # 2500
data.test  <- data[ind3,] # 2500

```

## Visualizations {#sec-viz}

### Summary Statistics

Summary Statistics for Categorical variables are as follows:

```{r}
#Summary Statistics of the Categorical Data#
knit_print.skim_df <- function(x, ...) {
  cat("Skim summary:\n")
  cat("-------------\n")
  gt_obj <- gt(x) %>%
    tab_header(
      title = "Skim summary"
    )
  # Return the gt table
  return(gt_obj)
}

my_skim <- skim_with(base=sfl(n=length),factor=sfl(ordered=NULL), numeric=sfl(p0=NULL,p100=NULL,hist = NULL))
summary_cat <- my_skim(data.train)
knit_print(summary_cat)
```

Summary for Numerical Data is given in @tbl-summary-numerical

```{r}
#| label: tbl-summary-numerical
#| tbl-cap: Summary statistics of the Numerical Data
summary_age <- data.train %>%
  dplyr::summarise('Variables'="age",
            'Mean' = mean(age),
            'Median' = median(age),
            'St.Dev' = sd(age),
            'Min' = min(age),
            'Max' = max(age),
            'IQR' = quantile(age,0.75)-quantile(age,0.25),
            'Sample_size' = n())
summary_campaign <- data.train %>%
  dplyr::summarise('Variables'="campaign",
            'Mean' = mean(campaign),
            'Median' = median(campaign),
            'St.Dev' = sd(campaign),
            'Min' = min(campaign),
            'Max' = max(campaign),
            'IQR' = quantile(campaign,0.75)-quantile(campaign,0.25),
            'Sample_size' = n())
summary_previous <- data.train %>%
  dplyr::summarise('Variables'="previous",
            'Mean' = mean(previous),
            'Median' = median(previous),
            'St.Dev' = sd(previous),
            'Min' = min(previous),
            'Max' = max(previous),
            'IQR' = quantile(previous,0.75)-quantile(previous,0.25),
            'Sample_size' = n())
summary_emp.var.rate <- data.train %>%
  dplyr::summarise('Variables'="emp.var.rate",
            'Mean' = mean(emp.var.rate),
            'Median' = median(emp.var.rate),
            'St.Dev' = sd(emp.var.rate),
            'Min' = min(emp.var.rate),
            'Max' = max(emp.var.rate),
            'IQR' = quantile(emp.var.rate,0.75)-quantile(emp.var.rate,0.25),
            'Sample_size' = n())
summary_cons.price.idx <- data.train %>%
  dplyr::summarise('Variables'="cons.price.idx",
            'Mean' = mean(cons.price.idx),
            'Median' = median(cons.price.idx),
            'St.Dev' = sd(cons.price.idx),
            'Min' = min(cons.price.idx),
            'Max' = max(cons.price.idx),
            'IQR' = quantile(cons.price.idx,0.75)-quantile(cons.price.idx,0.25),
            'Sample_size' = n())
summary_cons.conf.idx <- data.train %>%
  dplyr::summarise('Variables'="cons.conf.idx",
            'Mean' = mean(cons.conf.idx),
            'Median' = median(cons.conf.idx),
            'St.Dev' = sd(cons.conf.idx),
            'Min' = min(cons.conf.idx),
            'Max' = max(cons.conf.idx),
            'IQR' = quantile(cons.conf.idx,0.75)-quantile(cons.conf.idx,0.25),
            'Sample_size' = n())
summary_euribor3m <- data.train %>%
  dplyr::summarise('Variables'="euribor3m",
            'Mean' = mean(euribor3m),
            'Median' = median(euribor3m),
            'St.Dev' = sd(euribor3m),
            'Min' = min(euribor3m),
            'Max' = max(euribor3m),
            'IQR' = quantile(euribor3m,0.75)-quantile(euribor3m,0.25),
            'Sample_size' = n())
summary_nr.employed <- data.train %>%
  dplyr::summarise('Variables'="nr.employed",
            'Mean' = mean(nr.employed),
            'Median' = median(nr.employed),
            'St.Dev' = sd(nr.employed),
            'Min' = min(nr.employed),
            'Max' = max(nr.employed),
            'IQR' = quantile(nr.employed,0.75)-quantile(nr.employed,0.25),
            'Sample_size' = n())
combined_summary <- bind_rows(summary_age, summary_campaign, summary_previous, summary_emp.var.rate, summary_cons.price.idx, summary_cons.conf.idx, summary_euribor3m,
                              summary_nr.employed)
combined_summary |>
  gt() |>
  fmt_number(decimals=2) |>
  cols_label(
    Variables=html("Variables"),
    Mean = html("Mean"),
    Median = html("Median"),
    St.Dev = html("Std. Dev"),
    Min = html("Min"),
    Max = html("Max"),
    IQR = html("IQR"),
    Sample_size = html("Sample Size")
  ) 
```

In the summary statistics above, it is evident that the continuous
variables exhibit significant variability, with notable differences
observed in dispersion of their minimum and maximum values.

### Histograms

The histograms to observe the data structures for all the variables are
constructed.It can be observed:

-   **job**: More number of customers have jobs in 'admin' followed by
    'blue-collar'

-   **education**: The maximum number of customers have a university
    degree. There are almost negligible 'illerterate' customers (only 2)

-   **marital**: The marital status of maximum number of customers is
    married

-   **housing**: More customers have a housing loan, but the difference
    compared to those without is not significant.

-   **loan**: Most customers do not have a personal loan but few
    customers do.

-   **contact**: More customers were contacted by the type 'cellular'

-   **month**: May was month of the year in which the most customers
    were contacted last

-   **poutcome**: The outcome of the previous marketing campaign is
    'nonexistent' for the most customers

It can be observed that the response variable 'y' is quite imbalanced.

```{r}
#Visualizations# UNIVARIATE PLOTS
#Job#
job_dist <- data.train %>%
  group_by(job) %>%
  dplyr::summarise(job_count = n()) %>%
  arrange(desc(job_count))
job_plot <- ggplot(data = job_dist, aes(x = reorder(job, -job_count), y = job_count)) +
  geom_col() + 
  labs(title = "Job Distribution of Customers",
       x = "Job Level",
       y = "Count")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
#EDUCATION#
edu <- data.train %>%
  group_by(education) %>%
  dplyr::summarise(edu_count = n()) %>%
  arrange(desc(edu_count))
edu_plot <- ggplot(data = edu, aes(x = reorder(education, -edu_count), y = edu_count)) +
  geom_col() + # aes(fill = education)
  labs(title = "Education Distribution of Customers",
       x = "Education Level",
       y = "Count")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))# maximum have university degree followed by high school
#MARITAL STATUS#
marital_dist <- data.train %>%
  group_by(marital) %>%
  dplyr::summarise(marital_count = n()) %>%
  arrange(desc(marital_count))
marital_plot <- ggplot(data = marital_dist, aes(x = reorder(marital, -marital_count), y = marital_count)) +
  geom_col() +
  labs(title = "Marital Distribution of Customers",
       x = "Marital Status",
       y = "Count")+
  theme_bw() #maximum are married
#HOUSING#
housing_dist <- data.train %>%
  group_by(housing) %>%
  dplyr::summarise(housing_count = n()) %>%
  arrange(desc(housing_count))
housing_plot <- ggplot(data = housing_dist, aes(x = reorder(housing, -housing_count), y = housing_count)) +
  geom_col() +
  labs(title = "Housing Loan Distribution of Customers",
       x = "Housing",
       y = "Count")+
  theme_bw() # not a big difference
#LOAN#
loan_dist <- data.train %>%
  group_by(loan) %>%
  dplyr::summarise(loan_count = n()) %>%
  arrange(desc(loan_count))
loan_plot <- ggplot(data = loan_dist, aes(x = reorder(loan, -loan_count), y = loan_count)) +
  geom_col() +
  labs(title = "Personal Loan Distribution of Customers",
       x = "Loan",
       y = "Count")+
  theme_bw() # not a big difference
#contact#
contact_dist <- data.train %>%
  group_by(contact) %>%
  dplyr::summarise(contact_count = n()) %>%
  arrange(desc(contact_count))
contact_plot <- ggplot(data = contact_dist, aes(x = reorder(contact, -contact_count), y = contact_count)) +
  geom_col() +
  labs(title = "Contact Distribution of Customers",
       x = "Contact",
       y = "Count")+
  theme_bw()  #Maximum have cellular
#MONTH#
month_dist <- data.train %>%
  group_by(month) %>%
  dplyr::summarise(month_count = n()) %>%
  arrange(desc(month_count))
month_plot <- ggplot(data = month_dist, aes(x = reorder(month, -month_count), y = month_count)) +
  geom_col() +
  labs(title = "Monthly Distribution of Customers",
       x = "Month",
       y = "Count")+
  theme_bw() #May is the highest then July, least is December
#poutcome#
poutcome_dist <- data.train %>%
  group_by(poutcome) %>%
  dplyr::summarise(poutcome_count = n()) %>%
  arrange(desc(poutcome_count))
poutcome_plot <- ggplot(data = poutcome_dist, aes(x = reorder(poutcome, -poutcome_count), y = poutcome_count)) +
  geom_col() +
  labs(title = "poutcome Distribution of Customers",
       x = "poutcome",
       y = "Count")+
  theme_bw() #maximum are NON EXISTANT
#Response Y#
y_dist <- data.train %>%
  group_by(y) %>%
  dplyr::summarise(y_count = n()) %>%
  arrange(desc(y_count))
y_plot <- ggplot(data = y_dist, aes(x = reorder(y, -y_count), y = y_count)) +
  geom_col(aes(fill = y)) +
  labs(title = "y Distribution of Customers",
       x = "y",
       y = "Count")+
  theme_bw()
grid.arrange(job_plot, edu_plot, marital_plot, housing_plot, loan_plot, contact_plot, month_plot,  poutcome_plot, y_plot,ncol = 4)
```

```{r}
# Bivariate Plot
# Job Distribution with respect to y
job_y_plot <- ggplot(data.train, aes(x = job, fill = y)) +
  geom_bar(position = "stack") +
  labs(title = "Job Distribution of Customers by y",
       x = "Job",
       y = "Count",
       fill = "y")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Education Distribution with respect to y
edu_y_plot <- ggplot(data.train, aes(x = education, fill = y)) +
  geom_bar(position = "stack") +
  labs(title = "Education Distribution of Customers by y",
       x = "Education Level",
       y = "Count",
       fill = "y")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Marital Distribution with respect to y
marital_y_plot <- ggplot(data.train, aes(x = marital, fill = y)) +
  geom_bar(position = "stack") +
  labs(title = "Marital Distribution of Customers by y",
       x = "Marital Status",
       y = "Count",
       fill = "y")+
  theme_bw()

# Housing Distribution with respect to y
housing_y_plot <- ggplot(data.train, aes(x = housing, fill = y)) +
  geom_bar(position = "stack") +
  labs(title = "Housing Distribution of Customers by y",
       x = "Housing",
       y = "Count",
       fill = "y")+
  theme_bw()
# Loan Distribution with respect to y
loan_y_plot <- ggplot(data.train, aes(x = loan, fill = y)) +
  geom_bar(position = "stack") +
  labs(title = "Loan Distribution of Customers by y",
       x = "Loan",
       y = "Count",
       fill = "y")+
  theme_bw()
# Contact Distribution with respect to y
contact_y_plot <- ggplot(data.train, aes(x = contact, fill = y)) +
  geom_bar(position = "stack") +
  labs(title = "Contact Distribution of Customers by y",
       x = "Contact",
       y = "Count",
       fill = "y")+
  theme_bw()
# Month Distribution with respect to y
month_y_plot <- ggplot(data.train, aes(x = month, fill = y)) +
  geom_bar(position = "stack") +
  labs(title = "Month Distribution of Customers by y",
       x = "Month",
       y = "Count",
       fill = "y")+
  theme_bw()
# poutcome Distribution with respect to y
poutcome_y_plot <- ggplot(data.train, aes(x = poutcome, fill = y)) +
  geom_bar(position = "stack") +
  labs(title = "Poutcome Distribution of Customers by y",
       x = "Poutcome",
       y = "Count",
       fill = "y")+
  theme_bw()
grid.arrange(job_y_plot, edu_y_plot, marital_y_plot, housing_y_plot, loan_y_plot,contact_y_plot, month_y_plot, poutcome_y_plot, ncol = 4)
```

### Correlation using Scatterplot Matrix

```{r}
#Visualization for NUMERIC DATA#
df_new_numeric <- df_new %>%
  select_if(is.numeric)
ggpairs(df_new_numeric,columns=1:8, ggplot2::aes(colour = y, alpha = 0.2), title = "Scatterplot matrix")+
  theme_bw()
```

The above displays the correlation between different continuous
variables as per the 'yes;' and 'no' of response variable 'y'. It can be
observed there are mostly weak correlations. However, strong correlation
can be noticed between euribor3m and emp.var.rate, nr.employed &
emp.var.rate , euribor3m & nr.employed. Moderate correlation can be
observed between cons.price.idx & emp.var.rate, euribor3m &
cons.price.idx.

### Test for Independence

To test if there is a significant association between the two
categorical variables, chi-square test and fisher tests were conducted.
For both the tests, the null hypothesis assumes that the variables are
independent, while the alternative hypothesis suggests otherwise. All
variables have p \< 0.05 except 'loan' and 'housing. This suggests that
there is no significant association between the two categorical
variables.

```{r}
# Checking Significance
chi_job <- chisq.test(data.train$job, data.train$y) # p-value < 0.05
chi_marital <- chisq.test(data.train$marital, data.train$y) # p-value < 0.05
chi_edu <- chisq.test(data.train$education, data.train$y) # Chi-squared approximation may be incorrect
chi_housing <- chisq.test(data.train$housing, data.train$y) # p-value > 0.05
chi_loan <- chisq.test(data.train$loan, data.train$y) # p-value > 0.05
chi_contact <- chisq.test(data.train$contact, data.train$y) # p-value < 0.05
chi_month <- chisq.test(data.train$month, data.train$y) # Chi-squared approximation may be incorrect
chi_poutcome <- chisq.test(data.train$poutcome, data.train$y) # p-value < 0.05

# education
fisger_edu <-fisher.test(table(data.train$education, data.train$y), simulate.p.value = TRUE) # p-value < 0.05

# month 
fisher_month <- fisher.test(table(data.train$month, data.train$y), simulate.p.value = TRUE) # p-value < 0.05
```

### Outliers

The data was further explored to check if there are any outliers in the
continuous data. Few outlier were observed in age.Logarithmic
transformation has already account for most, however age had few
outliers. In 'cons.price.idx' columns, 53 observation were outliers but
for the values -26.9. Since the outliers are few, it is decided to keep
them in the dataset.

```{r}
# OUTLIERS AND boxplots OF CONTINUOUS VARIABLES#
b1 <- ggplot(data = data.train, aes(x = y, y = age)) +
  geom_boxplot(aes(fill = y)) +  
  labs(x = "y", y = "Age")

# emp.var.rate
b2 <- ggplot(data.train, aes(x = y, y = emp.var.rate)) +
  geom_boxplot(aes(fill = y)) +  
  labs(x = "y", y = "emp.var.rate")

# cons.price.idx
b3 <- ggplot(data.train, aes(x = y, y = cons.price.idx)) +
  geom_boxplot(aes(fill = y)) +  
  labs(x = "y", y = "cons.price.idx")

# cons.conf.idx
b4 <- ggplot(data.train, aes(x = y, y = cons.conf.idx)) +
  geom_boxplot(aes(fill = y)) +  
  labs(x = "y", y = "cons.conf.idx")

# euribor3m
b5 <- ggplot(data.train, aes(x = y, y = euribor3m)) +
  geom_boxplot(aes(fill = y)) +  
  labs(x = "y", y = "euribor3m")

# nr.employed
b6 <- ggplot(data.train, aes(x = y, y = nr.employed)) +
  geom_boxplot(aes(fill = y)) +  
  labs(x = "y", y = "nr.employed")

age_outliers <- data.train %>%
  mutate(is_outlier = age > quantile(age, 0.75) + 1.5 * IQR(age) |
           age < quantile(age, 0.25) - 1.5 * IQR(age)) %>%
  filter(is_outlier) %>%
  ungroup() # 9 outliers

cons.conf.idx_outliers <- data.train %>%
  mutate(is_outlier = cons.conf.idx > quantile(cons.conf.idx, 0.75) + 1.5 * IQR(cons.conf.idx) |
           cons.conf.idx < quantile(cons.conf.idx, 0.25) - 1.5 * IQR(cons.conf.idx)) %>%
  filter(is_outlier) %>%
  ungroup() # -26.9 is the outlier value of 53 observations
```

As housing and loan were found to be insignificant, they are omitted
from the data sets.

```{r}
data.train <- data.train%>%
  dplyr::select(-housing,-loan)
data.test <- data.test%>%
  dplyr::select(-housing,-loan)
data.valid <- data.valid%>%
  dplyr::select(-housing,-loan)
```

To conduct a statistical analysis for a data set containing both
categorical and continuous variables, one-hot encoding is applied to
categorical variables. In this method, the categorical variables are
assigned a binary value of 1 or 0 so as to achieve a new column for each
category of each categorical variable.

```{r}
categorical_columns <- c(3:8)
# Perform one-hot encoding on selected columns using column numbers
encoded_data.train <- model.matrix(~ . - 1, data = data.train[, categorical_columns])
encoded_data.test <- model.matrix(~ . - 1, data = data.test[, categorical_columns])
encoded_data.valid <- model.matrix(~ . - 1, data = data.valid[, categorical_columns])

banking_data_train <- cbind(data.train[, 1:2], encoded_data.train, data.train[, setdiff(names(data.train), c(names(data.train)[1:2], names(data.train[, categorical_columns])))])
banking_data_test <- cbind(data.test[, 1:2], encoded_data.test, data.test[, setdiff(names(data.test), c(names(data.test)[1:2], names(data.test[, categorical_columns])))])
banking_data_valid <- cbind(data.valid[, 1:2], encoded_data.valid, data.valid[, setdiff(names(data.valid), c(names(data.valid)[1:2], names(data.valid[, categorical_columns])))])
```

However, as SVM and Tree based classification methods are compatible
with categorical variables, one-hot encoding will not be applied to it.

```{r}
banking_data_train_svm <- data.train
banking_data_test_svm <- data.test
banking_data_valid_svm <- data.valid

# Remove 'y_coded' column from the datasets
banking_data_train_svm <- banking_data_train_svm[, !(names(banking_data_train_svm) %in% 'y_coded')]
banking_data_test_svm <- banking_data_test_svm[, !(names(banking_data_test_svm) %in% 'y_coded')]
banking_data_valid_svm <- banking_data_valid_svm[, !(names(banking_data_valid_svm) %in% 'y_coded')]

banking_data_train_tree <- data.train
banking_data_test_tree <- data.test
banking_data_valid_tree <- data.valid

# Remove 'y_coded' column from the datasets
banking_data_train_tree <- banking_data_train_tree[, !(names(banking_data_train_tree) %in% 'y_coded')]
banking_data_test_tree <- banking_data_test_tree[, !(names(banking_data_test_tree) %in% 'y_coded')]
banking_data_valid_tree <- banking_data_valid_tree[, !(names(banking_data_valid_tree) %in% 'y_coded')]

# Setting the factor levels for 'y' in each dataset
banking_data_train_tree$y <- factor(banking_data_train_tree$y, levels = c("yes", "no"))
banking_data_test_tree$y <- factor(banking_data_test_tree$y, levels = c("yes", "no"))
banking_data_valid_tree$y <- factor(banking_data_valid_tree$y, levels = c("yes", "no"))
```

# Classification Methods

### k-Nearest Neighbours {#sec-knn}

k-Nearest Neighbours is a popular classification method in machine
learning. It learns from the training data by remembering the features
of each point, calculating the distance between points, choosing optimal
k and finally makes predictions on validation data by predicting the
class as the class label of majority of the k nearest points
(neighbours). The method's effectiveness is evaluated using measures
such as F1 score, sensitivity, and classification rates.

An interesting point worth noticing is that knn does not assume a
parametric model for the underlying data. This characterisitc
facilitates easy understanding of forecasts, particularly for data such
as our bank marketing dataset. It is utilised in this study to
categorise the client subscription for term deposits. The method's
capacity for pattern recognition is another reason for its application.
Given the training on various factors, its principle that similar data
points belong to the same class is useful for forecasting whether or not
a client would subscribe to a term deposit.

```{r}
#Numerical summary for features in training data

skim(banking_data_train)
#Features have large ranges and sd
#Diff range and sd in features
#Euclidean  distance dominated by these features
#Remedy: Standardise
```

It has been previously observed in @sec-EDA that the features in the
data have different ranges and standard deviations. Since features with
wide ranges and standard deviations might dominate the computation,
adopting Euclidean Distance raises certain challenges. To deal with
this, the features have been standardized so that they have zero mean
and a variance of one.

```{r}
#Checking the column names
names(banking_data_train)

#Mean of features
var.mean<-apply(banking_data_train[ ,c(3:41)],2,mean)

#Standard Deviation of features
var.sd<-apply(banking_data_train[ ,c(3:41)],2,sd)

#Standardizing training, validation and test sets
banking_data_train_scale <-t(apply(banking_data_train[,c(3:41)],1,function(x) (x-var.mean)/var.sd))
banking_data_valid_scale <-t(apply(banking_data_valid[,c(3:41)],1,function(x) (x-var.mean)/var.sd))
banking_data_test_scale <-t(apply(banking_data_test[,c(3:41)],1,function(x) (x-var.mean)/var.sd))
```

Moving forward the most important decision to be made when predicting
classes using knn is the choice of k. Given that the response variable
is binary, ideal choice should be an odd value for k to avoid ties in
categories. To begin with, the focus is on evaluating the correct
classification rate on the validation set for different values of k.

```{r}
#Finding optimal value of k
set.seed(1)
K <- c(1:25) 
valid.corr <- c()
for (k in K){
  valid.pred <- knn(banking_data_train_scale,banking_data_valid_scale,banking_data_train[,1],k=k)
  valid.corr[k] <- mean(banking_data_train[,1] == valid.pred)
}

```

@fig-kopt plots the validation accuracy against different values of k to
visually analyse the significance of k on model performance. The optimal
k value would be the one which gives the highest correct classification
rate, in this case, k=25 is the optimal choice. This step also makes
predictions on the validation set which is utilised in assessing the
performance of knn as a classification method.

```{r}
#| eval: true
#| label: fig-kopt
#| fig-cap: Correct Classification rates on the validation set for different values of k
set.seed(1)
#Plotting validation correct classification rate
plot(K, valid.corr, type="b", ylab="validation correct classification rate")

```

```{r}
#Assigning optimal k to k.opt
set.seed(1)
k.opt<-(which.max(valid.corr))
k.opt # 23
```

```{r}
#Checking levels of valid.pred
levels_valid_pred <- levels(factor(valid.pred,levels=c("yes","no")))
levels_valid_pred

#Checking levels of data.valid$y
levels_data_valid_y <- levels(factor(banking_data_valid$y,levels=c("yes","no")))
levels_data_valid_y

#Confusion matrix and Statistics
conf_matrix <- confusionMatrix(reference =factor(banking_data_valid$y,levels=levels_data_valid_y),
                               data = factor(valid.pred, levels = levels_valid_pred))$table
conf_matrix

#Extracting important values
TP<-46
FP<-22
FN<-251
TN<-2181

#Accuracy
accuracy<-(46+2181)/(46+2181+251+22)
accuracy

#Precision
precision<-46/(46+22)
precision

#Recall
recall<-46/(46+251)
recall

#F1 score
f1<-(2*(precision*recall))/(precision+recall)
f1

```

Finally, to formally analyse the performance the Confusion Matrix in
@tbl-confmat is considered. It provides a comparison between actual
classification and predicted classification on validation set making
ground for obtaining informative metrics.

knn correctly classifies 46 instances where it predicted "yes"
(subscription) and the clients indeed subscribed (True Positives).

It incorrectly classified 22 instances as "yes" (subscription) when the
clients did not subscribe. These are cases where knn incorrectly
identified potential subscribers (False Positives).

It also incorrectly classified 251 instances as "no" (no subscription)
when the clients actually subscribed. These are cases where knn failed
to identify potential subscribers (False Negatives).

Lastly, knn correctly classified 2181 instances as "no" (no
subscription) where the clients did not subscribe (True Negatives).

In summary, the confusion matrix helps the bank understand the
performance of its subscription prediction model, enabling it to refine
its marketing strategies to maximize subscription rates.

Given these values, we further calculate performance metrics which
indicate the performance of knn as a classification method for bank
marketing data.

```{r}
#| label: tbl-confmat
#| tbl-cap: Confusion Matrix for knn

#Confusion matrix data with labels
conf<- matrix(c(TP, FP, FN, TN), nrow = 2, ncol = 2, 
               dimnames = list(c("Predicted.Yes", "Predicted.No"), c("Actual Yes", "Actual No")))

#Creating data frame with row and column labels
conf_matrix <- data.frame(conf)

#Confusion Matrix 
gt(conf_matrix,rownames_to_stub=T)
```

```{r}
#| label: tbl-metrics
#| tbl-cap: Performance Metrics
metrics<-as.data.frame(list(Metric = c("Accuracy", "Precision", "Recall","F1"),
                 Value = c(0.89,0.68,0.15,0.25)))

gt(metrics)
```

@tb-metrics shows high overall accuracy of 0.89, indicating it correctly
classified 89.08% of the cases in the validation set. This suggests the
model is, in general, good at distinguishing between "yes" and "no"
classes. A precision of 67.64% further indicates a moderate proportion
of positive predictions were actually positive. However, a low recall of
15.49% indicates missing substantial proportion of actually positive
cases. Overall, the low F1 score indicates even though the method avoids
many false positive, it misses many true positives. This indicate a weak
overall performance of knn on the bank marketing data.

### Linear Discriminant Analysis {#sec-lda}

Linear Discriminant Analysis is a classification method used to define
linear relationships between features. It aims to find a linear
combination of features that best classify and differentiates between
the classes in data. It assumes that the features are normally
distributed and begins by modelling the features, x, in classes G=g.
Following that, LDA uses Bayes' theorem to estimate the probability
Pr(G=g\|X=x). We thus utilise this probability to determine the category
based on a threshold for classification of data points.

The most important step in LDA is checking the validity of assumptions
berfore actually applying LDA. First, classes are mutually exclusive and
exhaustive. Second, the prior probability that a randomly chosen class
belongs to gth class is $$\pi_g=Pr(G=g)$$. Third and the most important,
features follow a multivariate Gaussian distribution. Fourth, which acts
as a deciding factor as to whether we should use LDA or QDA is
class-covraince matrix. We would use LDA in case of equal covariance and
QDA in case of unequal covraiance.

In the context of our bank marketing data, although the first two
assumptions seem theoretically valid from \@@sec-EDA, we observed in
@scat in @sec-EDA that very few variables like age, cons.price.idx and
cons.conf.idx fairly follow a Gaussian distribution while others do not.
This violates an essential assumption of LDA. To further analyse the
validity of LDA we examine the class-covariance matrix as variance
between groups in features. In this case, we observe that large
difference in variance between groups indicating violation of another of
the key assumptions.

These violations thus deem LDA as an inappropriate choice for our bank
marketing dataset. However, it would be interesting to observe LDA on
three of the features which satisfy the Gaussian distribution. These
features are age, cons.price.idx and cons.conf.idx. Hence, we perform
LDA on these variables.

```{r}
set.seed(1)

#Computing variance for each group
for (i in c(3:41)){
  cat(colnames(banking_data_train[,c(3:41)])[i],sep="\n")
  print(aggregate((banking_data_train[,c(3:41)]),by=list(banking_data_train$y),var)) 
}

#ggpairs: fairly Gaussian distribution for age but not for other variables like nr.employed and cons.price.index
#grouped variance show large difference in variance between groups, like nr.employed and cons.price.index

#LDA model
banking.lda <- lda(y~age+cons.price.idx+cons.conf.idx,data=banking_data_train)
banking.lda 

```

```{r}
#| label: fig-lda
#| fig-cap: Linear Discriminant Analysis Histogram Per Class
#Plotting result from lda
set.seed(1)
plot(banking.lda)
```

In @fig-lda it can observed that the linear discriminant function hardly
separates the two classe. There is evidence of overlap. However, it
should be noted that only a subset of the many features of interest are
explored in this method. To further assess the performance of LDA
examine numerical summaries are examined.

```{r}
#Prediction 
lda.pred<-predict(banking.lda,banking_data_valid)

#Checking values
names(lda.pred)

#Checking class values
head(lda.pred$class,10)

#Error rate
acc<-mean(banking_data_valid$y==lda.pred$class)
1-acc #0.1188

#Classification rate for prediction on the validation set
lda.tab<-table(banking_data_valid$y,lda.pred$class)
lda.tab  

#class specific classification rates for prediction on the validation set
lda.tab.prop<-prop.table(lda.tab,margin=1)
lda.tab.prop

```

Initially, it is observed that approximately 11.88% of the predictions
made by the LDA model on the validation dataset were incorrect. On
further analysis of confusion matrix it is observed that out of 2203
clients who did not subscribe for the term deposit, LDA misclassifies
zero, which is very good. However, of 297 clients who said yes to
subscription, LDA incorrectly predicts all of them as no.

In another attempt to modify LDA for better results Bayes discrimination
rule which states to assign an object to the class with the largest
posterior probability, is utilised. Here, if a standard threshold of 50%
for posterior probability of a yes for subscription is set. A few
thresholds to analyse the deviations in the result are explored for
better results in the analysis.

```{r}
#Adjusting the classification threshold as 25% 
data.pred.new1<-ifelse(lda.pred$posterior[ ,2]>0.25,"yes","no") 

#Adjusting the classification threshold as 50% 
data.pred.new2<-ifelse(lda.pred$posterior[ ,2]>0.50,"yes","no") 

#Adjusting the classification threshold as 75% 
data.pred.new3<-ifelse(lda.pred$posterior[ ,2]>0.75,"yes","no") 


#Error rate for 25%
acc1<-mean(banking_data_valid$y==data.pred.new1)
1-acc1 #0.1268

#Error rate for 50%
acc2<-mean(banking_data_valid$y==data.pred.new2)
1-acc2 #0.1188

#Error rate for 75%
acc3<-mean(banking_data_valid$y==data.pred.new3)
1-acc3 #0.1188

table.predict<-table(banking_data_valid$y,data.pred.new2)
table.predict

table.rate<-prop.table(table.predict,margin=1)
table.rate

```

Using 25% as the threshold, it is observed that the error rate goes up
to 12.68% of the predictions made by the LDA model. This makes the LDA
prediction worse. For a threshold of 50% and75%, the error rate stays
the same as earlier indicating no improvement in the model predictions
with confusion matrix also indicating no improvement.

A final look at the ROC curve and AUC would help in solidifying the
arguments even further.

```{r}
set.seed(1)
#ROC and AUC
prediction<-prediction(lda.pred$posterior[ ,2],banking_data_valid$y)
perf<-performance(prediction,"tpr","fpr")
plot(perf,main="ROC")

#To calculate AUC
AUC<-performance(prediction,"auc")
AUC@y.values[[1]]
```

The ROC is far from indicating a good classification techniques, it is
fairly closer to a random guess. The AUC for the data subset being
considered is closer to random guess, 0.59, hence providing evidence of
a not so good classifier.

In conclusion, it is important to state that this LDA analysis was
conducted only on a subset of features which fairly followed a
multivariate Gaussian distribution. Hence, it cannot be considered as an
overall classifier for the subscription responses since other features
we have been observed in @sec-EDA to have had a significant impact on
the response variable weren't eligible for LDA. Even for the few
variables of interest, the results don't seem very impressive indicating
LDA as a not a very good classifier for the bank marketing data.

### Quadratic Discriminant Analysis {#sec-qda}

Quadratic Discriminant Analysis (QDA) is an extension of Linear
Discriminant Analysis which models categorical variables, specially
those having non-linear decision boundary and unequal variance in
grouped features. is a statistical classification technique used for
modeling and predicting categorical outcomes. QDA models the
distribution of each class for predictions.

Similar to LDA, QDA assumes that each class comes from a distinct
multivariate normal distribution. Unlike LDA, which assumes a single
common covariance matrix for all classes, QDA gives the flexibility for
each class to have its own coveriance matrix in or to capture complex
decision boundaries.

QDA is actually a better choice for the bank marekting dataset as it
considers the case with unequal class covariance matrix which is the
case in the data being analysed in this study. So, in addtion to LDA,
QDA is also performed on the few fairly normally distributed features in
order to compare the two methods.

```{r}
#LDA model
set.seed(1)
banking.qda <- qda(y~age+cons.price.idx+cons.conf.idx,data=banking_data_train)
banking.qda 

qda.pred<- predict(banking.qda,banking_data_valid)
```

To begin with, the class are predicted on validation data and will be
considering the performance of age, cons.price.idx and cons.conf.idx.
However, it should be noted that only a subset of the many features
opposed to other important features of interest to be explored are
utilised, due to the violation of normality assumption. Hence, to assess
the performance of QDA the numerical summaries are examined.

```{r}
#Checking values
names(qda.pred)

#Checking class values
head(lda.pred$class,10)

acc<-mean(banking_data_valid$y == qda.pred$class)
1-acc # 0.1256

#Classification rate for prediction on the validation set
qda.tab<-table(banking_data_valid$y,qda.pred$class)
qda.tab  

#class specific classification rates for prediction on the validation set
qda.tab.prop<-prop.table(qda.tab,margin=1)
qda.tab.prop

```

Initially, it is observed that approximately 12.56% of the predictions
made by the QDA model on the validation dataset were incorrect, Which is
higher as compared to LDA. On further analysis of confusion matrix it is
observed that out of 2203 clients who did not subscribe for the term
deposit, QDA misclassifies 82 (3.72%), which implies poor results as
compared to LDA which misclassified zero. However, out of 297 clients
who said yes to subscription, QDA incorrectly predicts 232 (78.11%) of
them as no, which indicates better results than LDA.

In another attempt to modify QDA for better results, Bayes
discrimination rule which states we assign an object to the class with
the largest posterior probability, that is class which has the largest
QDF, is utilised. Here, a few thresholds to analyse the deviations in
the results are explored.

```{r}
#Adjusting the classification threshold as 25% 
qda.pred.new1<-ifelse(qda.pred$posterior[ ,2]>0.25,"yes","no") 

#Adjusting the classification threshold as 50% 
qda.pred.new2<-ifelse(qda.pred$posterior[ ,2]>0.50,"yes","no") 

#Adjusting the classification threshold as 75% 
qda.pred.new3<-ifelse(qda.pred$posterior[ ,2]>0.75,"yes","no") 


#Error rate for 25%
acc1<-mean(banking_data_valid$y==qda.pred.new1)
1-acc1 #0.1264

#Error rate for 50%
acc2<-mean(banking_data_valid$y==qda.pred.new2)
1-acc2 #0.1256

#Error rate for 75%
acc3<-mean(banking_data_valid$y==qda.pred.new3)
1-acc3 #0.1248

table.predict<-table(banking_data_valid$y,data.pred.new3)
table.predict

table.rate<-prop.table(table.predict,margin=1)
table.rate

```

Using 25% as the threshold, it is observed that the error rate goes down
slightly to 12.64% indicating a minor improvement from LDA with the same
threshold. For a threshold of 50% and 75%, the error rate imporoves on
increasing the threshold indicating improvement in the model
predictions. However at these thresholds, LDA had a slighty lower error
rate. In contrast, the confusion matrix of QDA now indicates the same
results as LDA.

A final look at the ROC curve and AUC would help in solidifying the
arguments even further.

```{r}
set.seed(1)
#ROC and AUC
prediction<-prediction(qda.pred$posterior[ ,2],banking_data_valid$y)
perf<-performance(prediction,"tpr","fpr")
plot(perf,main="ROC for QDA")

#To calculate AUC
AUC<-performance(prediction,"auc")
AUC@y.values[[1]]
```

The ROC again is far from indicating a good classification technique, it
indicates slightly better results for QDA as compared to LDA. The AUC
for the data subset being considered is closer to random guess, 0.67 but
still better than LDA (0.59). Hence providing evidence of QDA being a
better classifier relative to LDA, however not so good a classifier in
general.

In conclusion, it is extremely important to consider that LDA and QDA
was conducted only on a subset of features which fairly followed a
multivariate Gaussian distribution. Hence, it cannot be considered as an
overall classifier for the subscription responses since other features
we have observed in @sec-EDA to have had a significant impact on the
repsonse variable weren't eligible for LDA and QDA. Even for the few
variables of interest, the results don't seem very impressive. Although
QDA does a slightly better job than LDA, both of them are not very good
classifiers for the whole data in general as the assumption of normality
is violated in the data.

## Support Vector Machines {#sec-Method_3}

The purpose of an Support vector machines (SVMs) is to fit a hyperplane
that divides the points of the two classes into distinct zones. By
kernel trick, it can be mapped to a new space to solve the
classification problem of the original space: the resulting function is
linear in new space, but non-linear in the primitive. This method is
intended for the binary classification, which is suitable for this data
set. And it is compatible to deal with the categorical variables, so
there is no requirement to convert them to numerical variables.

The choices of kernel functions are linear kernel, polynomial kernel,
radial basis function (RBF) kernel and sigmoid kernel, the first three
are chosen. And the classification for different values of parameters is
fitted, the one with the smallest validation prediction error is finally
chosen for each function.

-   The model fulfills the assumptions

-   'type = C-classification' is used since the aim is to classify

-   All functions specifies the cost parameter

-   Polynomial kernel adds a parameter D:degree, when degree \> 1, we
    are fitting a hyperplane in a higher-dimensional space involving
    polynomials of degree

-   In RBF kernel, gamma is positive constant which defines the
    influence of a single sample when mapping

```{r, include=FALSE}
# Scale training set
banking_data_train_svm_numeric <- banking_data_train_svm %>%
  select_if(is.numeric)
banking_data_train_svm_not_numeric <- banking_data_train_svm %>% 
  select_if(Negate(is.numeric))
num_min <- sapply(banking_data_train_svm_numeric, min)
num_max <- sapply(banking_data_train_svm_numeric, max)
banking_data_train_svm_numeric <- (banking_data_train_svm_numeric - num_min) / (num_max - num_min)
banking_data_train_svm <- cbind(banking_data_train_svm_numeric, banking_data_train_svm_not_numeric)
```

```{r, include=FALSE}
# Scale Validation set
banking_data_valid_svm_numeric <- banking_data_valid_svm %>%
  select_if(is.numeric)
banking_data_valid_svm_not_numeric <- banking_data_valid_svm %>% 
  select_if(Negate(is.numeric))
num_min <- sapply(banking_data_valid_svm_numeric, min)
num_max <- sapply(banking_data_valid_svm_numeric, max)
banking_data_valid_svm_numeric <- (banking_data_valid_svm_numeric - num_min) / (num_max - num_min)
banking_data_valid_svm <- cbind(banking_data_valid_svm_numeric, banking_data_valid_svm_not_numeric)
```

```{r, include=FALSE}
# Scale Testing set
banking_data_test_svm_numeric <- banking_data_test_svm %>%
  select_if(is.numeric)
banking_data_test_svm_not_numeric <- banking_data_test_svm %>% 
  select_if(Negate(is.numeric))
num_min <- sapply(banking_data_test_svm_numeric, min)
num_max <- sapply(banking_data_test_svm_numeric, max)
banking_data_test_svm_numeric <- (banking_data_test_svm_numeric - num_min) / (num_max - num_min)
banking_data_test_svm <- cbind(banking_data_test_svm_numeric, banking_data_test_svm_not_numeric)
```

**Assumptions**

-   As mentioned earlier, SVM requires the response variable to be a
    binary variable.

    This assumption is valid due to construction and interpretation of
    the data set

-   The characteristics should ideally be on a similar scale.

    The process of scaling has been applied on banking_data_train_svm,
    banking_data_valid_svm and banking_data_test_svm to achieve this.

-   The performance of support vector machines depends crucially on the
    hyperparameters chosen.

    For each function, parameters are carefully chosen by by comparing
    them and minimizing validation prediction error.

-   For linear boundaries, as specified by the by linear kernel, after
    the model is fit, there's a good Accuracy(0.889) for linear kernel
    but low F1 Score(0.283), indicating that there is not enough
    evidence if the data is linearly se

    So soft margin is used, and also turn to non-linear using kernel
    trick

```{r, include=FALSE}
# some functions will be used for comparison
pred.error<-function(pred,truth){
mean(pred!=truth)
}
sensitivity <- function(table) {
  TP <- table["yes", "yes"]
  FN <- table["no", "yes"]
  sensitivity <- TP / (TP + FN)
  return(sensitivity)
}
specificity <- function(table) {
  TN <- table["no", "no"]
  FP <- table["yes", "no"]
  specificity <- TN / (TN + FP)
  return(specificity)
}
accuracy <- function(table) {
  TP <- table["yes", "yes"]
  TN <- table["no", "no"]
  FP <- table["yes", "no"]
  FN <- table["no", "yes"]
  accuracy <- (TP + TN) / (TP + TN + FP + FN)
  return(accuracy)
}
f1_score <- function(table) {
  TP <- table["yes", "yes"]
  FP <- table["yes", "no"]
  FN <- table["no", "yes"]
  precision <- TP / (TP + FP)
  recall <- TP / (TP + FN)
  f1_score <- 2 * (precision * recall) / (precision + recall)
  return(f1_score)
}
```

**linear kernel**

```{r, include=FALSE}
# cost
# range is based on Lab
# 100 is too large for iteration
C.val1 <- c(0.01,0.1,1,10)
C.error1 <- numeric(length(C.val1))

for(i in 1:length(C.val1)){
  model <- svm(y~., data = banking_data_train_svm, type="C-classification", kernel="linear", cost = C.val1[i])
    pred.model <- predict(model, banking_data_valid_svm)
    C.error1[i] <- pred.error(pred.model, banking_data_valid_svm$y)
}
C.error1
C.sel1 <- C.val1[min(which.min(C.error1))]
# 0.1

plot(C.val1, C.error1, type="b")
abline(v=C.sel1, lty=2)

final.svm1 <- svm(y~., data=banking_data_train_svm, kernel="linear", cost=C.sel1, type="C-classification")
summary(final.svm1)
# 1194 support vectors to create decision boundary
# 1194/5000=23.88% of the training data

pred.test <- predict(final.svm1, banking_data_test_svm)
pred.error(pred.test, banking_data_test_svm$y)

table1 <- table(banking_data_train_svm$y, predict(final.svm1))
table1

sensitivity_1 <- sensitivity(table1)
specificity_1 <- specificity(table1)
accuracy_1 <- accuracy(table1)
f1_score_1 <- f1_score(table1)
```

**polynomial kernel**

```{r, include=FALSE}
# cost
# degree
# range is based on Lab
# 100 is too large for iteration
C.val2 <- c(0.01,0.1,1,10)
D.val2 <- 2:5
C.error2 = matrix(nrow = length(C.val2), ncol = length(D.val2))

for(i in 1:length(C.val2)){
  for(j in 1:length(D.val2)){
    model <- svm(y~., data = banking_data_train_svm, type="C-classification", kernel="polynomial", cost = C.val2[i], degree = D.val2[j])
    pred.model <- predict(model, banking_data_valid_svm)
    C.error2[i,j] = pred.error(pred.model, banking_data_valid_svm$y)
  }
}
C.error2
min.index2 <- which.min(C.error2)
min_indices <- arrayInd(min.index2, .dim = dim(C.error2))
C.sel2 <- C.val2[min_indices[1]]
# 10
D.sel2 <- D.val2[min_indices[2]]
# 2

final.svm2 <- svm(y~., data = banking_data_train_svm, type="C-classification", kernel="polynomial", cost = C.sel2, degree = D.sel2)
summary(final.svm2)

pred.test <- predict(final.svm2, banking_data_test_svm)
pred.error(pred.test, banking_data_test_svm$y)

table2 <- table(banking_data_train_svm$y, predict(final.svm2))
table2
sensitivity_2 <- sensitivity(table2) # 0.68
specificity_2 <- specificity(table2) # 0.91
accuracy_2 <- accuracy(table2) # 0.8998
f1_score_2 <- f1_score(table2) # 0.2770
```

**radial kernel**

```{r, include=FALSE}
# cost
# gamma
# range is based on Lab
# 100 is too large for iteration
C.val3 <- c(0.01,0.1,1,10)
G.val3 <- c(0.001,0.01,0.1,1,10)
C.error3 = matrix(nrow = length(C.val3), ncol = length(G.val3))

for(i in 1:length(C.val3)){
  for(j in 1:length(G.val3)){
    model <- svm(y~., data = banking_data_train_svm, type="C-classification", kernel="radial", cost = C.val3[i], gamma = G.val3[j])
    pred.model <- predict(model, banking_data_valid_svm)
    C.error3[i,j] = pred.error(pred.model, banking_data_valid_svm$y)
  }
}
C.error3
min.index3 <- which.min(C.error3)
min_indices <- arrayInd(min.index3, .dim = dim(C.error3))
C.sel3 <- C.val3[min_indices[1]]
# 10
G.sel3 <- G.val3[min_indices[2]]
# 0.01
final.svm3 <- svm(y~., data = banking_data_train_svm, type="C-classification", kernel="radial", cost = C.sel3, gamma = G.sel3)
summary(final.svm3)

pred.test <- predict(final.svm3, banking_data_test_svm)
pred.error(pred.test, banking_data_test_svm$y)

table3 <- table(banking_data_train_svm$y, predict(final.svm3))
table3
sensitivity_3 <- sensitivity(table3)
specificity_3 <- specificity(table3)
accuracy_3 <- accuracy(table3)
f1_score_3 <- f1_score(table3)
```

```{r}
#| label: tbl-SVM
#| tbl-cap: "different kernels' performance"
table <- data.frame(
  Metric = c("Sensitivity", "Specificity", "Accuracy", "F1 Score"),
  Model_1 = c(sensitivity_1, specificity_1, accuracy_1, f1_score_1),
  Model_2 = c(sensitivity_2, specificity_2, accuracy_2, f1_score_2),
  Model_3 = c(sensitivity_3, specificity_3, accuracy_3, f1_score_3)
)
knitr::kable(table, digits = 3)
```

According to @tbl-SVM, on the basis of Sensitivity, it is better to
choose the polynomial kernel. But if F1 score is set to be used to
compare with other models, then models with linear kernel or radial
kernel may be chosen.

The linear kernel and radial kernel perform similarly in terms of
classification decision boundaries, possibly because the Radial Basis
Function (RBF) does not significantly improve performance. Since the
data doesn't need to be mapped into a higher dimensional space, the
linear kernel alone may be sufficient.

## Decision Trees {#sec-Method_4}

Decision trees are versatile supervised learning algorithms used for
classification and regression tasks, constructing hierarchical
structures that partition data based on feature values to predict
outcomes. Constructing an effective decision tree involves selecting
parameters like complexity, minimum splits, and depth, which are
fine-tuned using grid search techniques to optimize performance metrics
such as accuracy and F1 score. Additionally, tree pruning techniques,
such as minimum error reduction and smallest tree size, are applied to
prevent overfitting and simplify the model. Evaluating these models
using metrics like sensitivity, specificity, and AUC helps in
understanding their predictive power and generalizability, guiding the
selection of the most effective model configuration.

```{r}
evaluate_model <- function(model_name, data = banking_data_valid_tree, threshold = 0.5) {  # Added 'data' as an argument
  
  # Ensure actual outcomes are factored with levels in the order "yes", "no"
  actual <- factor(data$y, levels = c("yes", "no"))
  
  # Extract predicted probabilities and classify based on the threshold
  predicted_probs <- predict(get(deparse(substitute(model_name))), newdata = data, type = "prob")[, "yes"]
  predicted_classes <- ifelse(predicted_probs > threshold, "yes", "no")
  predicted_classes <- factor(predicted_classes, levels = c("yes", "no"))
  
  # Generate Confusion Matrix and calculate metrics
  cm <- confusionMatrix(predicted_classes, actual)
  accuracy <- cm$overall['Accuracy']
  sensitivity <- cm$byClass['Sensitivity']
  specificity <- cm$byClass['Specificity']
  F1_score <- cm$byClass['F1']
  
  # Calculate AUC using pROC
  ROC_curve <- roc(response = actual, predictor = as.numeric(predicted_probs), levels = rev(levels(actual)))
  AUC <- pROC::auc(ROC_curve)
  
  # Compile and return the results
  results_df <- data.frame(
    Metric = c("Accuracy", "Sensitivity", "Specificity", "F1 Score", "AUC"),
    Value = c(accuracy, sensitivity, specificity, F1_score, AUC)
  )
  
  print(results_df, row.names = FALSE)  # Print the results without row names
}

```

```{r}
print_cm <- function(model, data = banking_data_valid_tree, threshold = 0.5) {
  # Ensure required libraries are loaded
  if (!requireNamespace("caret", quietly = TRUE)) {
    install.packages("caret")
  }
  library(caret)
  
  # Extract actual outcomes and ensure the factor levels are in the correct order
  actual <- factor(data$y, levels = c("yes", "no"))  # Ensure 'y' matches your response variable name
  
  # Predict probabilities and classify based on the threshold
  predicted_probs <- predict(model, newdata = data, type = "prob")
  # Ensure the correct index for 'yes' class; it assumes that 'yes' is the second level
  # Modify if your 'yes' class is at a different index
  predicted_classes <- ifelse(predicted_probs[, "yes"] > threshold, "yes", "no") 
  predicted_classes <- factor(predicted_classes, levels = c("yes", "no"))
  
  # Create the confusion matrix
  cm <- confusionMatrix(predicted_classes, actual)
  
  # Print the confusion matrix
  print(cm)  # Only print the table part for brevity; remove '$table' if full output is needed
}

```

```{r}
evaluate_model_summary <- function(model, model_name, data = banking_data_valid_tree, threshold = 0.5) {

  actual <- factor(data$y, levels = c("yes", "no"))  # Ensure actual outcomes are factored correctly
  predicted_probs <- predict(model, newdata = data, type = "prob")[, "yes"]  # Assume 'yes' is the second level
  predicted_classes <- ifelse(predicted_probs > threshold, "yes", "no")
  predicted_classes <- factor(predicted_classes, levels = c("yes", "no"))  # Ensure predicted classes are factored correctly
  
  # Calculate metrics
  cm <- confusionMatrix(predicted_classes, actual)
  accuracy <- cm$overall['Accuracy']
  sensitivity <- cm$byClass['Sensitivity']
  specificity <- cm$byClass['Specificity']
  F1_score <- cm$byClass['F1']
  
  # AUC
  auc_value <- pROC::auc(roc(response = actual, predictor = as.numeric(predicted_probs), levels = rev(levels(actual))))
  
  # Return the formatted summary
  return(data.frame(Model = model_name, Accuracy = accuracy, Sensitivity = sensitivity, 
                    Specificity = specificity, F1_Score = F1_score, AUC = auc_value))
}

```

```{r}
visualize_metrics <- function(model, data = banking_data_valid_tree, thresholds = seq(0, 1, by = 0.01)) {
  
  # Initialize vectors to store metrics
  accuracy <- sensitivity <- specificity <- F1_scores <- numeric(length(thresholds))
  
  # Extract actual outcomes and convert to factor with correct levels
  actual <- factor(data$y, levels = c("yes", "no"))
  
  # Predict probabilities for the positive class
  predicted_probs <- predict(model, newdata = data, type = "prob")[, "yes"]

  # Compute metrics for each threshold
  for (i in seq_along(thresholds)) {
    threshold <- thresholds[i]
    predicted_classes <- ifelse(predicted_probs > threshold, "yes", "no")
    predicted_classes <- factor(predicted_classes, levels = c("yes", "no"))
    
    cm <- confusionMatrix(predicted_classes, actual)
    accuracy[i] <- cm$overall['Accuracy']
    sensitivity[i] <- cm$byClass['Sensitivity']
    specificity[i] <- cm$byClass['Specificity']
    F1_scores[i] <- cm$byClass['F1']
  }

  # Compute AUC separately since it's not threshold-dependent
  AUC_value <- pROC::auc(roc(actual, predicted_probs))
  
  # Combine the metrics into a data frame for plotting
  results <- data.frame(
    Threshold = thresholds, 
    Accuracy = accuracy, 
    Sensitivity = sensitivity, 
    Specificity = specificity, 
    F1_Score = F1_scores, 
    AUC = rep(AUC_value, length(thresholds))
  )
  
  # Plot the metrics against thresholds
  ggplot(results, aes(x = Threshold)) + 
    geom_line(aes(y = Accuracy, colour = "Accuracy"), size = 1.2) +
    geom_line(aes(y = Sensitivity, colour = "Sensitivity"), size = 1.2) +
    geom_line(aes(y = Specificity, colour = "Specificity"), size = 1.2) +
    geom_line(aes(y = F1_Score, colour = "F1 Score"), size = 1.2) +
    geom_line(aes(y = AUC, colour = "AUC"), size = 1.2, linetype = "dashed") +
    scale_colour_manual(values = c(
      Accuracy = "blue", Sensitivity = "red", 
      Specificity = "green", "F1 Score" = "purple", AUC = "orange")) +
    labs(y = "Metric Value", title = "Model Performance Metrics Across Different Thresholds") +
    theme_minimal()
}

```

### Model Development

A base decision tree was constructed using the rpart package in R. This
base model served as a foundation for comparing subsequent, more refined
models. Following the development of the base decision tree model,
parameter tuning is conducted to enhance its performance. The tuning
process involves exploring a grid of parameters, also refereed as
stopping criteria:

-   cp (Complexity Parameter)

-   minsplit (Minimum Split)

-   minbucket (Minimum Bucket)

-   maxdepth (Maximum Depth)

For each combination of these parameters, a new decision tree model was
trained. The models' performances were evaluated, allowing to identify
the optimal set of parameters based on our predefined metrics,
particularly focusing on improving the F1 score due to its relevance in
scenarios with imbalanced classes.

```{r}
# Build base tree model
model_tree <- rpart(y ~., data = banking_data_train_tree, method = "class")
```

```{r}
# Define your parameter grid
set.seed(1)
params_grid <- expand.grid(cp = seq(0.01, 0.1, by = 0.01), 
                           minsplit = c(10, 20, 30, 40),
                           minbucket = c(5, 10, 15, 20),
                           maxdepth = c(5, 10, 15, 20))

results <- data.frame(cp = numeric(), minsplit = numeric(), minbucket = numeric(), 
                      maxdepth = numeric(), Accuracy = numeric(), F1 = numeric())

for(i in 1:nrow(params_grid)) {
  params <- params_grid[i, ]
  
  model <- rpart(y ~ ., data = banking_data_train_tree, method = "class",
                 control = rpart.control(cp = params$cp, 
                                         minsplit = params$minsplit,
                                         minbucket = params$minbucket,
                                         maxdepth = params$maxdepth))
  
  predictions <- predict(model, banking_data_valid_tree, type = "class")
  # Ensure predictions are factors with the same levels as the training outcomes
  predictions <- factor(predictions, levels = levels(banking_data_valid_tree$y))
  
  cm <- confusionMatrix(predictions, banking_data_valid_tree$y)
  accuracy <- cm$overall['Accuracy']
  f1 <- cm$byClass['F1']
  
  # Create a temporary data frame to store the iteration results
  temp_results <- data.frame(cp = params$cp, minsplit = params$minsplit, 
                             minbucket = params$minbucket, maxdepth = params$maxdepth, 
                             Accuracy = accuracy, F1 = f1)
  # Combine with main results data frame
  results <- rbind(results, temp_results)
}

# Identify the best-performing set of parameters
best_params <- results[which.max(results$Accuracy), ]
print(best_params)
```

```{r}
model_tree_criteria <- rpart(y ~ ., data = banking_data_train_tree, method = "class", cp = 0.01, minsplit = 10, minbucket = 15, maxdepth = 5)
```

### Tree Pruning

In addition to tuning the base model, tree pruning strategies are
implemented to reduce overfitting and simplify the model. Two approaches
are utilized:

-   **Minimum Error Strategy**: Pruning the tree based on a criterion
    that seeks to minimize prediction error, which could involve cutting
    back the tree until the additional complexity does not significantly
    reduce the model's error rate on validation data.

-   **Smallest Tree Strategy**: This approach involves reducing the tree
    to its simplest form without a significant drop in predictive
    accuracy, aiming for the most interpretable model.

```{r}
# Build full-grown tree by setting cp = -1
set.seed(1)
model_tree_full <- rpart(y ~ ., data = banking_data_train_tree, method = "class", cp = -1)
```

```{r}
## Pruning with Minimum Error
min_error_cp <- model_tree_full$cptable[which.min(model_tree_full$cptable[, "xerror"]), "CP"]
pruned_tree_min_error <- prune(model_tree_full, cp = min_error_cp)
```

```{r}
# Pruning with Smallest Tree
min_error <- min(model_tree_full$cptable[, "xerror"])
min_error_std <- model_tree_full$cptable[which.min(model_tree_full$cptable[, "xerror"]), "xstd"]
smallest_tree_cp <- model_tree_full$cptable[model_tree_full$cptable[, "xerror"] <= min_error + min_error_std, "CP"][1]
pruned_tree_smallest <- prune(model_tree_full, cp = smallest_tree_cp)
```

### Results

```{r}
# Apply function to each model
results_1 <- rbind(
  evaluate_model_summary(model_tree, "Base Tree"),
  evaluate_model_summary(model_tree_criteria, "Base Tree Tuned "),
  evaluate_model_summary(pruned_tree_min_error, "Pruned Tree Min Error"),
  evaluate_model_summary(pruned_tree_smallest, "Pruned Tree Smallest")
)

# View the results
results_table_1 <- results_1 %>%
  gt(rownames_to_stub = FALSE)%>%
  fmt_number(
    columns = everything(), 
    decimals = 3
  )
results_table_1
```

The performance metrics table for the decision tree models shows that
tuning and pruning strategies have minimal impact on the overall
accuracy, which remains consistent across all models. However, there is
a noticeable decline in sensitivity (recall) and F1 score as the models
move from the base tree to the pruned versions, indicating a potential
loss in the ability to correctly identify positive cases. Specifically,
both pruned models show identical decreases in sensitivity and F1 score
compared to the base models, suggesting that pruning may overly simplify
the model, reducing its effectiveness in identifying true positives
without significantly improving specificity or AUC. The marginal
increase in specificity and AUC in the tuned and pruned models does not
compensate for the loss in sensitivity and F1 score. Thus, while efforts
to optimize and simplify the models through tuning and pruning maintain
high specificity and slightly improved AUC, they may compromise the
models' ability to detect the positive class effectively.

### Interpretation

```{r}
rpart.plot(model_tree)
```

**Significance of Employment Rate (nr.employed)**: The tree splits first
on nr.employed, suggesting that the employment rate is a significant
predictor. Specifically, clients with nr.employed less than 5088 are
more likely to subscribe (the left branch), indicating that lower
employment rates might correlate with higher subscription rates,
possibly due to more aggressive investment or savings behaviors during
times of lower employment rates.

**Impact of Previous Outcomes (poutcome**): For clients within the lower
employment rate segment, the tree further distinguishes based on the
outcome of previous marketing campaigns (poutcome = success). Clients
who were successfully reached in previous campaigns are more likely to
subscribe (30% chance), highlighting the importance of previous positive
interactions with the bank.

**Influence of Contact Month (month)**: On the higher employment rate
side, the tree identifies the month of contact as a crucial factor,
particularly March and October (month = mar,oct). Clients contacted in
these months are more inclined to subscribe compared to other months,
possibly due to seasonal factors or end-of-financial-period
considerations.

**Role of Economic Conditions (euribor3m)**: For clients contacted
outside of March and October in conditions of higher employment rates,
the Euribor 3-month rate (euribor3m \>= 1.6) serves as a further
determinant. A higher Euribor rate seems to be associated with lower
subscription rates, likely reflecting broader economic conditions that
affect client decisions.

### Conclusion

The tree based models provide a comprehensive view of the factors
influencing term deposit subscriptions. The best performing model
highlights that variables such as the employment rate (nr.employed), the
outcome of previous marketing campaigns (poutcome), the month of contact
(month), and economic conditions (euribor3m) are pivotal in determining
client decisions. Despite these insights, the basic tree model exhibits
low sensitivity and F1 scores, indicating challenges in effectively
identifying true positives. However, the consistency in accuracy and
specificity across different tree models indicates a strong ability to
recognize true negatives.

## Bagging and Random Forest

In advancing the ensemble methods, bagging was initially employed to
reduce variance and enhance model robustness. Building on this, the
Random Forest method was applied, incorporating feature selection
randomness at each split. This added layer of randomness aimed to reduce
over-fitting further and improve model generalization. Parameters within
the Random Forest, including the number of trees and the maximum number
of features considered at each split, were systematically tuned based on
model performance metrics.

### Bagging

A bagging technique was utilized to reduce variance and enhance
robustness by generating multiple trees from different data set samples.
The model was then fine-tuned to improve the F1-score, thereby
optimizing the balance between precision and recall, highlighting the
effectiveness of ensemble methods in predictive modeling.

```{r}
library(randomForest)
set.seed(1)  # for reproducibility

# Perform Bagging (simulate by Random Forest)
bagging_model <- randomForest(y ~ ., data = banking_data_train_tree, mtry = ncol(banking_data_train_tree) - 1, ntree = 200)

```

```{r}
# Tuning Bagging Model
# Define the parameter grid without sample size
set.seed(1)
params_grid <- expand.grid(ntree = c(200, 500),
                           nodesize = c(1, 5),
                           stringsAsFactors = FALSE)  # Keep non-list parameters

# Initialize an empty data frame for tuning results
tuning_results <- data.frame(ntree = integer(), nodesize = integer(),
                             F1_Score = numeric(), stringsAsFactors = FALSE)


# Loop through the parameter grid
for (i in 1:nrow(params_grid)) {
  # Extract parameters
  params <- params_grid[i, ]
  
  # Train the model with specified parameters
  set.seed(1)  # For reproducibility
  model <- randomForest(y ~ ., data = banking_data_train_tree, mtry = ncol(banking_data_train_tree) - 1,
                        ntree = params$ntree, nodesize = params$nodesize)
  
  # Predict and evaluate
  predictions <- predict(model, banking_data_valid_tree, type = "class")
  cm <- confusionMatrix(predictions, banking_data_valid_tree$y)
  
  # Store results
  tuning_results <- rbind(tuning_results, c(params$ntree, params$nodesize, cm$byClass['F1']))
}

# Fix column names after rbind
names(tuning_results) <- c("ntree", "nodesize", "F1_Score")

# Identify best parameters based on F1 score
best_params <- tuning_results[which.max(tuning_results$F1_Score), ]
#print(best_params)
```

```{r}
set.seed(1)
bagging_model_tuned <- randomForest(y ~ ., data = banking_data_train_tree, mtry = ncol(banking_data_train_tree) - 1, ntree = 200, nodesize = 5)
#evaluate_model(bagging_model_tuned)
```

### Random Forest

The Random Forest method builds on bagging by introducing randomness in
feature selection at each tree split, enhancing diversity and reducing
over-fitting. This process improves model generalization by preventing
reliance on any single feature, allowing for the capture of complex
relationships.

```{r}
set.seed(1)
# Perform Random Forest
random_forest_model <- randomForest(y ~ ., data = banking_data_train_tree, ntree = 200)
```

```{r}
# Tuning Random Forest
# Define the parameter grid
set.seed(1)
params_grid <- expand.grid(
  mtry = round(seq(1, sqrt(ncol(banking_data_train_tree) - 1), length.out = 5)),  # Assuming last column is the response
  nodesize = c(1, 5),
  stringsAsFactors = FALSE
)

tuning_results <- data.frame(
  mtry = integer(), 
  nodesize = integer(),
  F1_Score = numeric(),
  stringsAsFactors = FALSE
)


# Assuming banking_data_valid_tree is your validation set
for (i in 1:nrow(params_grid)) {
  # Extract parameters
  params <- params_grid[i, ]
  
  # Train the model with specified parameters
  set.seed(1)  # For reproducibility
  rf_model <- randomForest(y ~ ., data = banking_data_train_tree, mtry = params$mtry,
                           ntree = 200, nodesize = params$nodesize)
  
  # Make predictions on the validation set
  predictions <- predict(rf_model, banking_data_valid_tree, type = "class")
  
  # Evaluate performance using F1 score
  cm <- confusionMatrix(predictions, banking_data_valid_tree$y)
  f1_score <- cm$byClass['F1']
  
  # Store results
  tuning_results <- rbind(tuning_results, c(params$mtry, params$nodesize, f1_score))
}

# Fix column names after rbind (if necessary)
names(tuning_results) <- c("mtry", "nodesize", "F1_Score")

best_params <- tuning_results[which.max(tuning_results$F1_Score), ]
#print(best_params)
```

```{r}
# Build Random Forest with Best Params
set.seed(1)
random_forest_model_tuned <- randomForest(y ~ ., data = banking_data_train_tree, ntree = 200, mtry = 3, nodesize = 5)
#evaluate_model(random_forest_model_tuned)
```

### Results

Results showed that tuning improves both Tree Bagging and Random Forest
models slightly in terms of accuracy, specificity, and AUC. However,
sensitivity remains relatively low for all models, indicating a
challenge in correctly identifying positive cases. The Random Forest
Tuned model shows the best overall balance of metrics, suggesting it as
the best model among those evaluated.

```{r}
# Apply function to each model
results_2 <- rbind(
  evaluate_model_summary(bagging_model, "Tree Bagging"),
  evaluate_model_summary(bagging_model_tuned, "Tree Bagging Tuned"),
  evaluate_model_summary(random_forest_model, "Random Forest"),
  evaluate_model_summary(random_forest_model_tuned, "Random Forest Tuned") 

)
# View the results
results_table_2 <- results_2 %>%
  gt(rownames_to_stub = FALSE)%>%
  fmt_number(
    columns = everything(), 
    decimals = 3
  )
results_table_2
```

### Interpretation

```{r}
varImpPlot(random_forest_model_tuned)
```

**Top Features**: The variables nr.employed, euribor3m, age, job, and
education appear to be the most significant in influencing the model's
predictions. These features are likely driving the majority of the
predictive power of the model, reflecting economic conditions
(nr.employed and euribor3m), as well as demographic and job-related
factors (age, job, education).

**Economic Indicators**: The prominence of nr.employed and euribor3m
highlights the impact of economic factors on the decision to subscribe
to term deposits. This aligns with expectations as economic conditions
often play a significant role in financial decision-making.

**Demographic and Job-related Factors**: The significance of age, job,
and education underscores the importance of demographic and employment
characteristics in influencing a client's likelihood to subscribe,
suggesting targeted marketing strategies could be effective.

### Conclusion

The analysis reveals that the tuned Random Forest model outperforms its
counterparts in accuracy, specificity, and F1 score, highlighting its
efficiency in predicting outcomes while maintaining a strong balance
between precision and recall. The feature importance plot indicates that
'euribor3m' and 'nr.employed' are the most significant predictors,
suggesting economic factors are crucial in influencing the model's
decisions. Given these insights, the tuned Random Forest model emerges
as a robust tool for forecasting, especially in scenarios where economic
indicators play a pivotal role.

# Results {#sec-Result}

In the evaluation of machine learning models, the Random Forest model
outperformed KNN, SVM, and Decision Tree variants in our banking data
set analysis. It showcased the highest F1 score (0.348) and a
well-balanced sensitivity (0.242) and specificity (0.978). This led to
its selection as the optimal model. When applying this model to the test
set, the Random Forest model achieved an accuracy of 0.902, sensitivity
of 0.278, specificity of 0.980, an F1 score of 0.387, and AUC of 0.783,
confirming its robustness and effectiveness for the task.

```{r}
# Test Results #
test_results <- evaluate_model(random_forest_model_tuned, banking_data_test_tree)

test_results <- test_results %>%
  gt(rownames_to_stub = FALSE)%>%
  fmt_number(
    columns = everything(), 
    decimals = 3
  )
test_results
```

# Conclusions {#sec-Conc}

In our comparative study, KNN, LDA/QDA, SVM and Tree-based models are
employed to predict whether a client is subscribed to a term deposit.
The Random Forest model emerged as the most effective with the highest
F1 score, which is particularly significant given the imbalanced nature
of our dataset, showcasing a good balance between predictive metrics.

The feature importance analysis in our Random Forest model highlights
the impact of economic indicators, such as 'nr.employed' and
'euribor3m', and demographic factors like 'age', 'job', and 'education'
on bank deposit subscriptions. These findings suggest that both broader
economic conditions and personal attributes significantly influence
customer decisions, guiding targeted strategies for marketing bank
products.

The Random Forest model excels due to its ensemble approach, which
combines multiple decision trees to reduce variance and prevent
overfitting, leading to more reliable predictions. Its effectiveness is
further enhanced by its ability to handle diverse variable scales
without pre-processing, crucial for our dataset with its wide-ranging
economic and demographic factors.

# References {#sec-Ref}

1.  Harrison, O. (2018) *Machine Learning Basics with the K-Nearest
    Neighbors Algorithm.* [Online Article]. Available from:
    <https://towardsdatascience.com/machine-learning-basics-with-the-k-nearest-neighbors-algorithm-6a6e71d01761>.

2.  IBM. (2023) *K-Nearest Neighbors (KNN).* [Online Resource].
    Available from:
    <https://www.ibm.com/docs/en/db2oc?topic=procedures-k-nearest-neighbors-knn>.

3.  Skand, K. (2017) *kNN(k-Nearest Neighbour) Algorithm in R.* [Online
    Article]. Available from:
    <https://rstudio-pubs-static.s3.amazonaws.com/316172_a857ca788d1441f8be1bcd1e31f0e875.html>.

4.  Hsu, Chih-Wei, Chih-Chung Chang, and Chih-Jen Lin. (2016) *A
    Practical Guide to Support Vector Classification.* [Department of
    Computer Science, National Taiwan University]. Taipei 106, Taiwan,
    May 19. Available from: <http://www.csie.ntu.edu.tw/~cjlin>.

5.  Xiaochen Yang. (2024) *Data Mining and Machine Learning 2023-24.*
    [Online Resource]. Available from:
    [https://moodle.gla.ac.uk/course/](https://moodle.gla.ac.uk/course/view.php?id=38609 "https://moodle.gla.ac.uk/course/view.php?id=38609")

6.  Grammar correction assistance was provided by ChatGPT, an AI
    language model developed by OpenAI, which utilizes natural language
    processing techniques for context-aware suggestions.
